---
title: "Making Simultaneous Calls to Platform"
date: "2017-08-14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Asychronous Programming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Concurrency in the Civis R Client

Just like most functions in R, all functions in `civis` block.  This means
that each function in a program must complete before the next function runs.
For instance,

```{r, eval=FALSE}
nap <- function(seconds) {
    Sys.sleep(seconds)
}

start <- Sys.time()
nap(1)
nap(2)
nap(3)
end <- Sys.time()
print(end - start)
```

This program takes 6 seconds to complete, since it takes 1 second for the
first `nap`, 2 for the second and 3 for the last. This program is easy to
reason about because each function is sequentially executed. Usually, that
is how we want our programs to run.

There are some exceptions to this rule. Sequentially executing each function
might be inconvenient if each `nap` took 30 minutes instead of a few seconds.
In that case, we might like our program to perform all 3 naps simultaneously.
In the above example, running all 3 naps simultaneously would take 3 seconds
(the length of the longest nap) rather than 6 seconds.

As all function calls in `civis` block, `civis` relies on the mature
R ecosystem for parallel programming to enable multiple simultaneous
tasks. The three packages we introduce are `future`, `foreach`, and
`parallel` (included in base R). For all packages, simultaneous tasks are
enabled by starting each task in a separate R process. Examples for building
several models in parallel with different libraries are included below. The
libraries have strengths and weaknesses and choosing which library to use is
often a matter of preference.

It is important to note that when calling `civis` functions, the computation
required to complete the task takes place in Platform.  For instance, during
a call to `civis_ml`, Platform builds the model while your laptop waits
for the task to complete.  This means that you don't have to worry about
running out of memory or cpu cores on your laptop when training dozens of
models, or when scoring a model on a very large population. The task being
parallized in the code below is simply the task of waiting for Platform to
send results back to your laptop.


## Building Many Models with `future`

```{r, eval=FALSE}
library(future)
library(civis)

# Define a concurrent backend with enough processes so each function
# we want to run concurrently has its own process. Here we'll need at least 2.
plan("multiprocess", workers=10)

# Load data
data(iris)
data(airquality)
airquality <- airquality[!is.na(airquality$Ozone),]  # remove missing in dv

# Create a future for each model, using the special %<-% assignment operator.
# These futures are created immediately, kicking off the models.
air_model %<-% civis_ml(airquality, "Ozone", "gradient_boosting_regressor")
iris_model %<-% civis_ml(iris, "Species", "sparse_logistic")

# At this point, `air_model` has not finished training yet. That's okay,
# the program will just wait until `air_model` is done before printing it.
print("airquality R^2:")
print(air_model$metrics$metrics$r_squared)
print("iris ROC:")
print(iris_model$metrics$metrics$roc_auc)
```


## Building Many Models with `foreach`

```{r, eval=FALSE}
library(parallel)
library(doParallel)
library(foreach)
library(civis)

# Register a local cluster with enough processes so each function
# we want to run concurrently has its own process. Here we'll
# need at least 3, with 1 for each model_type in model_types.
cluster <- makeCluster(10)
registerDoParallel(cluster)

# Model types to build
model_types <- c("sparse_logistic",
                 "gradient_boosting_classifier",
                 "random_forest_classifier")

# Load data
data(iris)

# Listen for multiple models to complete concurrently
model_results <- foreach(model_type=iter(model_types), .packages='civis') %dopar% {
    civis_ml(iris, "Species", model_type)
}
stopCluster(cluster)
print("ROC Results")
lapply(model_results, function(result) result$metrics$metrics$roc_auc)
```

## Building Many Models with `mcparallel`

Note: `mcparallel` relies on forking and thus is not available on Windows.

```{r, eval=FALSE}
library(civis)
library(parallel)

# Model types to build
model_types <- c("sparse_logistic",
                 "gradient_boosting_classifier",
                 "random_forest_classifier")

# Load data
data(iris)

# Loop over all models in parallel with a max of 10 processes
model_results <- mclapply(model_types, function(model_type) {
  civis_ml(iris, "Species", model_type)
}, mc.cores=10)

# Wait for all models simultaneously
print("ROC Results")
lapply(model_results, function(result) result$metrics$metrics$roc_auc)
```

## Operating System / Environment Specific Errors

Differences in operating systems and R environments may cause errors for
some users of the parallel libraries listed above.  In particular,
`mclapply` does not work on Windows and may not work in RStudio on
certain operating systems.  `future` may require `plan(multisession)` on
certain operating systems.  If you encounter an error parallelizing
functions in `civis`, we recommend first trying more than one method
listed above.  While we will address errors specific to `civis` with
regards to parallel code, the technicalities of parallel libraries in
R across operating systems and enviroments prevent us from providing
more general support for issues regarding parallelized code in R.
