---
title: "Data Import and Export"
author: "Patrick Miller, Keith Ingersoll"
date: "2017-08-14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Civis IO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Passing Data Back and Forth

Often the simplest, but most useful operation when working with the
Platform is to move data in and out.  From the perspective of the R
client, we call moving data from the Platform to the local
machine *reading*.  Likewise, moving data from the local
machine to the Platform is called *writing*.

The `civis` client handles data imports and exports in two basic ways:

1. Moving data directly between the R workspace and the Platform (the most common use case).
2. Moving data between the Platform and local csv files (this is useful for large data that doesn't fit into memory).

Data can be stored on Platform in two places:

1. Amazon Redshift, a SQL database.
2. Amazon S3, also referred to as the 'files' endpoint.

Tables in Redshift are accessed and modified using SQL queries. Tables in Redshift 
can be easily shared and used in multiple workflows by multiple people. 
However, importing and exporting even small files on Redshift can be slow.

R objects and arbitrary files can be stored on Amazon S3, and are accessed 
using a numeric file id. Data frames are uploaded as CSVs for portability,
and arbitrary R objects are serialized using `saveRDS` for speed and efficiency. 

## Reading Data Into R From Platform

The main workhose for getting data from Platform is `read_civis`.
This function is designed to work similarly to the built in function
`read.csv`, returning a dataframe from a table in Platform. For more flexibility,
`read_civis` can download files from Redshift using an SQL query, or download a 
file from S3 ('the files endpoint') using a file id.

To read from a table in Platform, simply provide  the name of the schema,
table within the schema, and the database:

```{r eval=FALSE}
df <- read_civis("schema.tablename", database = "my-database")
```

For convenience, a default database can be set in the package options, and not specified
in further calls to any IO function. If there is only one database available, this 
database will automatically be used as the default.
In the examples that follow, we assume that a default database has been set.

```{r eval=FALSE}
options(civis.default_db = "my-database")
df <- read_civis("schema.tablename")
```

`read_civis` accepts SQL queries when more flexibility is needed. This is accomplished
by wrapping `sql(...)` around a string containing the query. With `read_civis`, 
queries are always read only, and always return a `data.frame`.

```{r eval=FALSE}
query <- "SELECT * FROM table JOIN other_table USING id WHERE var1 < 23"
df <- read_civis(sql(query))
```

Finally, `read_civis` accepts a file id as the first argument to read in files 
from S3 as data frames. IDs are obtained from `write_civis_file`. 

```{r, eval=FALSE}
data(iris)
id <- write_civis_file(iris)
df <- read_civis(id)
```

For maximum flexibility, `read_civis` accepts parameters from `read.csv` which
can be used to define data types when the defaults are not appropriate. 
For instance, when numbers should be read in as characters or when strings 
shouldn't be read in as factors.

```{r eval=FALSE}
query <- "SELECT * FROM table JOIN other_table USING id WHERE var1 < 23"
df <- read_civis(sql(query), colClasses = "character")
df2 <- read_civis(sql(query), as.is = TRUE)
```


## Uploading Data to a Database

The complement to reading data into the R workspace is writing data
to the Platform.  The function `write_civis` uploads data frames or csv files to
an Amazon Redshift database. The function `write_civis_file` uploads R objects and 
arbitrary files to Amazon S3 (the files endpoint).

When creating a new table, `write_civis` relies on Platform to determine
data types.  Distkeys and sortkeys can optionally be set to improve query performance.
Again, we set a default database in these examples for convenience.

```{r eval=FALSE}
options(civis.default_db = "my_database")
df <- data.frame(x = rnorm(100), y = rnorm(100), z = rnorm(100))
write_civis(df, tablename = "schema.tablename",
            distkey = "id", sortkey1 = "date", sortkey2 = "type")
```

By default, `write_civis` will fail if the table passed in `tablename`
already exists.  Optionally, `write_civis` can append to an existing
table.  It may also delete all rows and then append (truncate). If
specific datatypes are required, a table may first be created with a
SQL `CREATE TABLE` command and then data can be inserted with
`write_civis`.

```{r eval=FALSE}
write_civis(df, tablename = "schema.tablename", if_exists = "append")
write_civis(df, tablename = "schema.tablename", if_exists = "truncate")
```

If a csv file is saved to disk but not loaded in the R workspace,
`write_civis` will upload the csv to Platform without needing
first load the csv into RAM. This can save time when a file is large. 
Uploading a csv directly to Platform is done by simply passing the file name
and path to `write_civis` as the first argument:

```{r, eval=FALSE}
write_civis("~/path/to/my_data.csv", tablename="schema.tablename")
```

## Uploading Data to S3

Finally, `write_civis_file` uploads data frames, R objects and files to Amazon S3, which is also
referred to as the 'files endpoint.' Data frames are uploaded as CSVs. 
R objects saved to the files endpoint and are serialized using `saveRDS`.

Data frames and R objects can be loaded back into memory by passing the 
file id to `read_civis`, and an appropriate `using` argument. 

```{r, eval = FALSE}
# Upload a data frame
data(iris)
id <- write_civis_file(iris)
iris2 <- read_civis(id)

# Upload an arbitrary R object
farm <- list(chickens = 1, ducks = 4, pigs = 2, cows = 1)
id <- write_civis_file(farm)
farm2 <- read_civis(id, using = readRDS)

```

When passed a file name and path, `write_civis_file` will upload the file to S3 as-is. 
To read the file back into memory, an appropriate function to convert the
file to a data frame must be provided to the `using` argument of `read_civis`. 
For example, a JSON file can be read back into R using `jsonlite::fromJSON`.

```{r, eval = FALSE}
id <- write_civis_file("path/to/my_data.json")
read_civis(id, using = jsonlite::fromJSON)
```


## Downloading Large Data Sets from Platform.

Occasionally, a table may be too large to store in memory. `download_civis`
can be used in place of `read_civis` to download data straight to disk from Platform.

Like `read_civis`, `download_civis` can download files from Amazon Redshift by passing
`schema.tablename`, or `sql(...)` as the first argument. Files can be downloaded from
Amazon S3 by passing the file id to `download_civis`. 

```{r eval=FALSE}
query <- "SELECT * FROM table JOIN other_table USING id WHERE var1 < 23"
download_civis(sql(query), file = "path/to/my_file.csv")
download_civis("schema.tablename", file = "path/to/my_file.csv")

id <- write_civis_file(iris)
download_civis(id, file = "path/to/my_iris.rds")
```


## Running Queries on Platform

Arbitrary queries can be run on Redshift using `query_civis`, which returns the meta-data
of the query.

```{r eval=FALSE}
q_res <- query_civis("GRANT ALL ON schema.my_table TO GROUP admin")
```

Existing queries can be re-run by passing the query id to `query_civis`:

```{r, eval = FALSE}
id <- q_res$id
query_civis(id)
```

## Common Errors

#### Civis API key not properly set or has expired.
Often an improper API key will return an error like below:
```{r, eval=FALSE}
 Error in api_key() : 
  The environmental variable CIVIS_API_KEY is not set. Add this to your .Renviron or call Sys.setenv(CIVIS_API_KEY = '<api_key>') 
```

However, there may be cases where the errors are less straightforward.  It
is a good idea to test that API credentials are properly set with a simple
call such as `civis::users_list_me()`.  See the README to set
up API keys correctly.

#### Query does not return any results.
This may happen if a table is empty 
or when no rows match a `WHERE` statement. To fix, double check that
the query is correct or the table is not empty.

```{r, eval=FALSE}
read_civis(sql("SELECT * FROM schema.tablename WHERE 1 = 0"))
Error in download_script_results(run$script_id, run$run_id) : 
  Query produced no output. 
```

#### Database not set correctly.
For both `read_civis` and `write_civis`, the database must be set to the
correct, case sensitive name (not hostname) of
the database. 

```{r, eval=FALSE}
 Error in get_db(database) : 
  Argument database is NULL and options("civis.default_db") not set. Set this option using options(civis.default_db = "my_database") 
```

To see a complete list of database names, run:

```{r, eval=FALSE}
sapply(databases_list(), function(x) x$name)
```
